The \dealii Library, Version 9.6

### 2024-06-01 9.6/paper.tex
First draft of 9.6 release paper

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Maximilian Bergbauer,
    Marco Feder,
    Marc Fehling,
    Johannes Heinz,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells,
    Stefano Zampini
  },
  pdftitle={The deal.II Library, Version 9.5, 2023},
}

\title{The \dealii Library, Version 9.5}

@@

\author[4]{Maximilian~Bergbauer}
\affil[4]{Institute for Computational Mechanics,
Technical University of Munich,
Boltzmannstr.~15, 85748 Garching, Germany.
{\texttt{maximilian.bergbauer@tum.de}}}

\author[5]{Marco~Feder}
\affil[5]{SISSA,
International School for Advanced Studies,
Via Bonomea 265,
34136, Trieste, Italy.
{\texttt{marco.feder/luca.heltai@sissa.it}}}

\author[2]{Marc~Fehling}

\author[6]{Johannes~Heinz}
 \affil[6]{Institute of Mechanics and Mechatronics,
   TU Wien,
   Getreidemarkt 9, 1060 Vienna, Austria
   {\texttt{johannes.heinz@tuwien.ac.at}}}

\author[7]{Timo~Heister}
 \affil[7]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister@clemson.edu}}}

\author[5]{Luca~Heltai}

 \author[8]{Martin~Kronbichler}
 \affil[8]{Institute of Mathematics,
   University of Augsburg,
   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
   {\texttt{martin.kronbichler/peter.muench@uni-a.de}}}

\author[9]{Matthias~Maier}
\affil[9]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[8,10]{Peter Munch}
 \affil[10]{Institute of Material Systems Modeling,
 Helmholtz-Zentrum Hereon,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany}

\author[11]{Jean-Paul~Pelteret}
\affil[11]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

@@

\author[12]{David Wells}
\affil[12]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

\author[13]{Stefano Zampini}
\affil[13]{Extreme Computing Research Center, King Abdullah University of Science and Technology,
  23955-6900 Thuwal, Saudi Arabia.
  {\texttt{stefano.zampini@gmail.com}}}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.5.
\end{abstract}

@@

\dealii version 9.5.0 was released July 7, 2023.
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii software library version 9.5. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Substantial updates and extensions to \dealii{}'s interfaces
    to other libraries (see Section~\ref{sec:external}). This
    includes, in particular, the integration of \kokkos (Section~\ref{sec:kokkos});
    additions and updates to the \petsc and \trilinos
    interfaces
    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}, respectively).
  \item Uniform handling of nonlinear solver
  packages (Section~\ref{sec:nonlinear}) and a uniform
    way of defining callbacks used by external libraries (Section~\ref{sec:callbacks}).
  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf}).
  \item Advances in non-matching support (see Section~\ref{sec:nonmatching}).
  \item New features related to linear algebra (see Section~\ref{sec:lac}).
  \item C++ language modernization (see Section~\ref{sec:language}).
  \item Build-system modernization (see Section~\ref{sec:buildsystem}).
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item The new function \texttt{CellAccessor::as\_dof\_handler\_iterator()}
  simplifies the conversion from a \texttt{Cell\-Accessor} to a \texttt{DoF\-Cell\-Accessor}.
  The old way,
\begin{c++}
const auto cell_dof = typename DoFHandler<dim, spacedim>::
  active_cell_iterator(&dof_handler.get_triangulation(),
    cell->level(), cell->index(), &dof_handler);
\end{c++}
was lengthy and error-prone. In contrast, the following function is
substantially clearer and more concise:
\begin{c++}
const auto cell_dof = cell->as_dof_handler_iterator(dof_handler);
\end{c++}
\item Several functions used intensively during initialization of
  \dealii-based programs, such as the refinement of triangulations, the
  enumeration of degrees of freedom, the setup of global-coarsening multigrid
  algorithms, and several evaluation functions of the \texttt{MappingQ} class
  representing a polynomial mapping of quadrilateral and hexahedral, have been
  overhauled to run more quickly and sometimes also consume less memory. These
  and related improvements are guided by several performance tests that are
  used to monitor the performance of the library over time.
\end{itemize}
%
The changelog lists more than 120 other features and bugfixes.

@@

%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Updates to interfaces to other packages}\label{sec:external}

For many operations, \dealii{} relies on external libraries -- some of
these are optional, others are mandatory (such as \boost and, now,
\kokkos); a complete list of external dependencies is provided in
Section~\ref{sec:cite}. A substantial amount of work has gone into
overhauling and extending these interfaces for the current release, as
detailed in the following sub-sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Integration of \kokkos}\label{sec:kokkos}

\kokkos \cite{trott2022} is a C++ library that enables the creation of
performance portable applications for all major high-performance computing
platforms. It implements a programming model that allows developers to write
code that can efficiently run on diverse architectures. \kokkos provides
abstractions for both parallel execution of code and data management. It
supports a wide range of backend programming models, including CUDA, HIP, SYCL,
HPX, OpenMP, and C++ threads, and continues to evolve with the development of
new hardware and corresponding backend options.

\dealii{} has, for several releases already, used CUDA to offload some
operations onto GPUs. It has also had interfaces to CUDA-based linear
algebra libraries. Yet, the diversification of GPU platforms away from
a single vendor (Nvidia) has made it clear that we need a different
strategy to support what users want. As a consequence, \kokkos has
become a mandatory dependency of \dealii{} as part of the current
release; if it is not found on a given system during configuration
time, then the library will fall back on a copy of \kokkos stored in
the \texttt{bundled/} directory in the same way as we already
interface with \boost.

In the current release, \texttt{LinearAlgebra::distributed::Vector} and the 
\texttt{CUDAWrappers::Matrix\allowbreak{}Free} framework are using \kokkos. This allows them 
to work on all the architectures supported by \kokkos. In particular, the step-64 example
can now also be run on the CPU and doesn't require a GPU anymore. \\
The \kokkos backend used by \dealii is \texttt{Kokkos::DefaultExecutionSpace} which corresponds
to the highest available backend in the hierarchy device, host-parallel, and host-serial 
at the time Kokkos was configured.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}

With the current release, \dealii{} now supports solvers for nonlinear problems
provided by several different external packages:
(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.

All three of these classes have a very similar interface, except that they vary in the kind of
algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
the three external solvers with a unified interface. The user can either let
\dealii{} decide which of the packages to use (depending on the current availability of
the external packages) or specify it manually.

The following code snippet shows a complete example using the new class for applying a Newton solver
by automatically choosing one of the available packages:
\begin{c++}
using NLS = NonlinearSolverSelector<VectorType>;

NLS::AdditionalData additional_data(
  NLS::automatic, // other options: kinsol, nox, petsc_snes 
  NLS::newton);

NLS nonlinear_solver(additional_data, mpi_communicator);

solver.reinit_vector       = [&](VectorType &x) {/*...*/};    
solver.residual            = [&](const VectorType &src,
                                 VectorType       &dst) {/*...*/};
solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
solver.apply_jacobian      = [&](const VectorType &src,
                                 VectorType       &dst) {/*...*/};
solver.solve_with_jacobian = [&](const VectorType &src,
                                 VectorType       &dst,
                                 const double      tol) {/*...*/};

solver.solve(current_solution);    
\end{c++}

Note that while the selector class provides a unified interface and therefore allows to
easily switch between backends, there are cases when users need to use the underlying
wrapper classes, if, e.g., the functionality is only provided by one implementation.
Details of each of the implementations are discussed in subsequent sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Updates and additions to the \petsc wrappers}\label{sec:petsc}

The \dealii classes wrapping \petsc objects have been rewritten in
substantial ways
to support \petsc's System of Nonlinear Equations Solver \snes and the Ordinary Differential Equations (ODE)
solver \ts \cite{abhyankar2018petsc}.
First, we briefly describe the most important improvements to the existing
classes and then outline the newly designed interfaces to the nonlinear solvers, together
with a new interface to the communication module \petscsf in \petsc \cite{zhang2021petscsf}.

Vector and matrix classes of the \petsc wrappers have been extended with an additional constructor
that can wrap an already existing \petsc vector or matrix, respectively.
The \texttt{BlockVector} and
\texttt{BlockSparseMatrix} classes now internally use
\petsc nested objects, i.e. \texttt{VECNEST} and  \texttt{MATNEST} respectively.
We have added a new class \texttt{PETScWrappers::PreconditionShell} to support user-defined
preconditioning that can be simply customized as the following code
snippet shows:
\begin{c++}
PETScWrappers::PreconditionShell preconditioner(/*...*/);
preconditioner.vmult = [&](const VectorType &src,
                           VectorType       &dst) {/*...*/};

\end{c++}
The resulting object can be passed to \petsc and used within the nonlinear solver hierarchy.
See Section~\ref{sec:callbacks} for additional information on such kind of callbacks.

The \petsc~\snes sub-package solves systems of nonlinear equations of the form $F(x) = 0$.
The interface to \snes in the class \texttt{PETScWrappers::NonlinearSolver} has been modeled on the already existing interface
to the \kinsol solver from the \sundials package: the nonlinear problem
is specified via a set of callbacks as shown above in Section~\ref{sec:nonlinear}.
%\begin{c++}
%PETScWrappers::NonlinearSolver<VectorType> solver(/*...*/);
%
%solver.residual            = [&](const VectorType &x,
%                                 VectorType       &F) {/*...*/};
%solver.setup_jacobian      = [&](const VectorType &x) {/*...*/};
%solver.solve_with_jacobian = [&](const VectorType &rhs,
%                                 VectorType       &sol) {/*...*/};
%solver.solve(x);
%\end{c++}
The default configuration is set up to use a Jacobian-free Newton-Krylov (JFNK)
approach~\cite{knoll2004jacobian}, using \texttt{solve\_with\_jacobian} as linear solver.
Numerous other solver configurations are possible and can be selected
programmatically using the constructor arguments,
or via the powerful command line customization of \petsc.
This includes, for example, Quasi-Newton methods, Anderson's acceleration, and
nonlinear preconditioning~\cite{brune2015composing}. The step-77
tutorial program has gained a section that discusses the use of these
new \snes interfaces.

The \petsc~\ts sub-package solves ODEs in explicit or implicit form \cite{abhyankar2018petsc}, i.e.:
\begin{eqnarray*}
%\frac{\partial u}{\partial t} = G(t,u) &\text{explicit},\\
%F(t,u,\frac{\partial u}{\partial t}) = 0 &\text{implicit}\\
\dot{u} = G(t,u) &\text{(explicit)},\\
F(t,u,\dot{u}) = 0 &\text{(implicit)}.
\end{eqnarray*}
The interface to \ts has been modeled on the already existing interfaces
to the \ida and \arkode solvers from the \sundials package. Specifically:
\begin{c++}
PETScWrappers::TimeDependentSolver<VectorType> solver(/*...*/);

// If solving udot = G(t,u)
solver.explicit_function = [&](const double     t,
                               const VectorType &u,
                               VectorType       &G) {/*...*/};
// If solving F(t,u,udot) = 0
solver.implicit_function = [&](const double     t,
                               const VectorType &u,
                               const VectorType &udot,
                               VectorType       &F) {/*...*/};
\end{c++}
In addition to specifying the function callbacks, users can further customize
the solution of the linearized equations $\alpha \frac{\partial{F}}{\partial{\dot{u}}} + \frac{\partial{F}}{\partial{u}}$
via the following callbacks:
\begin{c++}
solver.setup_jacobian = [&](const double     t,
                            const VectorType &u,
                            const VectorType &udot,
                            const double     alpha) {/*...*/};
solver.solve_with_jacobian = [&](const VectorType &rhs,
                                 VectorType       &sol) {/*...*/};
\end{c++}
As with \snes, the default configuration of an implicit solver is set up to
use a JFNK approach, and the entire suite of solvers offered by \petsc is available
programmatically or via the command line interface,
including adaptive time-stepping and Implicit-Explicit schemes.

We close this section by introducing the interface to the \petscsf sub-package,
the abstract communication model of \petsc. The \dealii interface to
\petscsf has been modeled on the existing \texttt{Utilities::MPI::Partitioner} and
\texttt{Utilities::MPI::NoncontiguousPartitioner} classes, with minimal changes for the
communication routines API; the equivalent classes based
on \petscsf are \texttt{PETScWrappers::Partitioner} and
\texttt{PETScWrappers::CommunicationPattern}.
Future developments will add support for GPU buffers.

We refer interested readers to our documentation for more advanced
functions of the \snes, \ts and \petscsf wrappers, and to the {\tt tests/petsc/} folder
for examples on how to use them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Interfaces to \trilinos{}' \texttt{Belos} and \texttt{NOX} packages}\label{sec:trilinos}

\trilinos is a large collection of individual sub-packages \cite{heroux2005trilinos,trilinos-web-page}. \dealii{}
has long had interfaces to the \trilinos packages that provide parallel
vector and matrix classes (both \texttt{Epetra} and \texttt{Tpetra}),
as well as a small number of linear algebra packages for iterative
solvers and preconditions. In the current release, there are now also
interfaces to two additional packages: \texttt{Belos} and \texttt{NOX}.

\texttt{Belos} is the
successor of the \trilinos package \texttt{AztecOO} and provides basic and advanced iterative solvers
that heavily rely on multivector operations. The interface of the
wrapper is similar to \dealii{}'s own iterative solvers:
\begin{c++}
TrilinosWrappers::SolverBelos<VectorType> solver(/*...*/);
solver.solve(matrix, x, r, preconditioner);
\end{c++}
The omitted constructor arguments allow selecting the actual iterative
solver to be used, along with other configuration options.

Secondly, we have added a wrapper to \texttt{NOX}, a
nonlinear solver library that is similar to both the \kinsol{} solver from
the \sundials{} package to which we already had interfaces, and also
to the \snes{} collections of functions from \petsc (see also
Section~\ref{sec:petsc}). 
%The interfaces to these three solvers are
%essentially the same and require setting function objects for
%computing the residual of the nonlinear problem, along with setting
%up, applying, and solving with the Jacobian matrix:
%\begin{c++}
%TrilinosWrappers::NOXSolver<VectorType> solver(/*...*/);
%
%solver.residual            = [ ](const VectorType &src,
%                                 VectorType       &dst) {/*...*/};
%solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
%solver.apply_jacobian      = [&](const VectorType &src,
%                                 VectorType       &dst) {/*...*/};
%solver.solve_with_jacobian = [&](const VectorType &src,
%                                 VectorType       &dst,
%                                 const double      tol) {/*...*/};
%
%solver.solve(solution);
%\end{c++}
%
As for the \snes{} interface above, the \texttt{NOX} interface is
drive by the callbacks shown in Section~\ref{sec:nonlinear}, and
implemented in the \texttt{TrilinosWrappers::NOXSolver} class.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Uniform error reporting in callbacks}\label{sec:callbacks}

With the current release, \dealii{} has gained interfaces to several
external packages that are largely driven via \textit{callbacks} --
see for example the code examples for nonlinear or ODE solvers in the previous sections.

This substantially enlarged use of callbacks used by different backend
libraries raises the issue that each underlying package has its own
convention on how success or error codes of these callbacks should be
encoded. In the case of \petsc{}'s \snes{} and \ts{}, and for \trilinos{}'s \texttt{NOX},
success is indicated by a zero integer return value, whereas failure is
indicated by a nonzero return value. On the other hand, the \sundials{}
packages indicate success by a zero integer return value, a
recoverable failure with a positive value, and an irrecoverable failure
with a negative value. Newer PETSc versions can deal with recoverable
failures as well, in some cases by calling back into PETSc from a
callback to set an error flag, in others by setting the elements of a
returned vector to NaN.
None of these conventions mesh well with C++
where error codes are generally indicated via exceptions.
It is conceivable that libraries we want to
interface with in the future use yet other conventions.

In order to shield those who write these callbacks from having to
learn the intricacies of the underlying libraries, we have adopted a
convention whereby user-provided callbacks are just regular functions
that return errors via exceptions as is common in C++. Internally, the
interfaces to different underlying libraries then translate these
exceptions into the appropriate error codes, saving the thrown exception for possible
later use; if an underlying library supports
recoverable errors, then a callback should indicate such an error by
throwing an object of the special type
\texttt{RecoverableUserCallbackError}. If one of the underlying
packages returns with an error caused by a user-provided callback
throwing an exception, then the wrapper code re-throws the previously
saved exception, allowing the calling site to catch it to obtain information
about what might have gone wrong.

This convention for user callbacks is documented in a glossary entry
that is also linked to from the documentation of all variables storing
callbacks.

@@

The current release includes numerous updates to the matrix-free
infrastructure, including:
\begin{itemize}
\item In release 9.3, we enabled parallel $hp$-operations in the matrix-free infrastructure.
The infrastructure did not work properly for cells
that do not have any degrees of freedom because they use \texttt{FE\_Nothing}. This has been
fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a means to enable
or disable cells, we have introduced the new class
\texttt{ElementActivationAndDeactivationMatrixFree}, which wraps a \texttt{MatrixFree} object, only loops over all
active cells, and optionally interprets faces between active and deactivated cells
as boundary faces. This functionality has enabled simulations in powder-bed-fusion additive
manufacturing in~\cite{proell2023highly}.
\item The matrix-free infrastructure allows interleaving cell loops with vector updates
  by providing \texttt{pre}/\texttt{post} functions that are run on index ranges.  The \dealii
  library uses this feature, e.g., to improve the performance of (preconditioned)
conjugate gradient solvers~\cite{kronbichler2022cg} as well as of relaxation and Chebyshev iterations (see Subsection~\ref{sec:lac}).
Up to release~9.3, the \texttt{pre}/\texttt{post} infrastructure was only supported for
continuous elements (cell loop); now, it also works for discontinuous
elements which also require face loops to assemble jump and penalty terms.
\item The operator \texttt{CellwiseInverseMassMatrix} now also efficiently
evaluates the inverse for coupling (dyadic) coefficients in the case of multiple
components:
\begin{align*}
\left(v_i, D_{ij} u_j  \right)_{\Omega^{(K)}}
\quad  1\le i,j \le c,
\end{align*}
with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ a tensorial
coefficient. The algorithm relies on the construction of the element mass matrix,
\begin{align*}
M =  ( I_1 \otimes N^T) (D \otimes I_2 ) ( I_1 \otimes N),
\end{align*}
with $N$ being the tabulated values of shape functions at quadrature and
$I_1,I_2$ identity matrices associated to $c$ vector components and the
quadrature points, respectively.
The algorithms assumes a square $N$:
\begin{align*}
M^{-1} =  ( I_1 \otimes N^{-1}) (D^{-1} \otimes I_2 ) ( I_1 \otimes N^{-T}).
\end{align*}
For hypercube-shaped cells, $N^{-1}$ has an explicit representation
again in terms of tensor products; for example, in 3d it can be
expressed as $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes
N_{1D}^{-1}$, allowing the use sum factorization \cite{kronbichler2016comparison}.
\end{itemize}

@@

The (matrix-free) non-matching support of \dealii heavily relies
on the classes \texttt{FE\-Point\-Eval\-u\-ation} and \texttt{RemotePointEvaluation},
which were introduced in release 9.3. While \texttt{FE\-Point\-Eval\-u\-a\-tion}
is responsible for efficient evaluation/integration at arbitrary (reference)
points within a cell, \texttt{RemotePointEvaluation} is responsible for
sorting points with regards to the cells they reside in, and for the
communication necessary to evaluate solutions on cells owned by other
MPI processes.

In the current release, we have considerably optimized \texttt{FEPointEvaluation}, e.g,
by caching the evaluated shape functions, templating loop bounds, and
exploiting the tensor-product structure of the shape functions if all points are
positioned on a face, a common use case in the context of fluid-structure
interaction. Furthermore, the extended class \texttt{NonMatching::MappingInfo}
allows for precomputing and storing metric terms, like the Jacobian, its determinant,
or the unit outer normal vectors. This is useful in cases in which these metric terms do not change
and can be reused, e.g., in the context of iterative solvers. This development
is part of an effort to make the interfaces of the
(matrix-free) non-matching support more similar to the ones of the
established matrix-free infrastructure of \dealii for fixed quadrature formulas.

In addition to these node-level performance optimizations, we added experimental
support for (i) generating intersections of distributed
non-matching grids and working on them,
and (ii) multigrid with non-nested levels. In the following, we describe these
in more detail.

\subsubsection{Intersected meshes}
In the current release, we added experimental support to compute intersections on \texttt{parallel::distributed::Triangulation} objects using \texttt{CGAL}~\cite{cgal-user-ref}.
For this purpose, we introduced a free function \texttt{distributed\_compute\_intersection\_locations()} that computes intersections and relevant information for communication from \texttt{intersection\_re\-quests}.
The data structure \texttt{intersection\_requests} is a vector indicating entities of a given triangulation that intersections are computed upon.
Each entity (face or cell) is described by a vector of vertices.
Currently, the function is placed in \texttt{GridTools::internal}, since the location and arguments of the function might still change in the future.
To compute intersections between two geometric entities, the function internally uses the new function \texttt{CGALWrappers::compute\_intersec\-tion\_of\_cells()}.

For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections to evaluate the underlying physical coupling terms.
The data structure returned by \texttt{distributed\_compute\_intersection\_locations()} can convert itself to a data structure that can be used to fill \texttt{RemotePointEvaluation}.
This conversion is triggered by the member function~\texttt{convert\_to\_distributed\_compute\_point\_locations\_internal()}, given the number of quadrature points per intersection.
The whole procedure is done without communication.
%
This functionality is useful since \texttt{RemotePointEvaluation} can now be used to access quantities at quadrature points on intersections without further ado.
To reduce the user's effort, we plan to add a wrapper that takes care of the described procedure and provides interfaces like \texttt{FEEvaluation} to access quantities easily, e.g., in a matrix-free loop.

The procedure described above and an early version of the wrapper have
been used successfully in~\cite{heinz2023high} to perform Nitsche-type
mortaring in the context of the conservative formulation of acoustic
equations discretized with a discontinuous Galerkin method to suppress
artificial modes.

\subsubsection{Non-nested multigrid}

\dealii has provided support for geometric multigrid methods (GMG)
for locally refined meshes nearly since its inception. (\dealii{} also
supports algebraic multigrid methods by interfacing to the external
libraries \petsc and \trilinos.)
Traditionally, these GMGs were based on local-smoothing
methods (described many years later in \cite{Kanschat2004,JanssenKanschat2011}, see
also~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
infrastructure, furthermore, allows globally coarsening the polynomial degree ($p$-multigrid),
with specific applicability to $hp$-adaptive methods.
In the current release, we have added support for the case
that multigrid levels are given by non-nested meshes~\cite{adams2002evaluation, bittencourt2001nonnested, bramble1991analysis}. An example for such meshes is
presented in Figure~\ref{fig:nonnested}.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \includegraphics[width=\textwidth]{png/mesh_0.png}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \includegraphics[width=\textwidth]{png/mesh_1.png}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \includegraphics[width=\textwidth]{png/mesh_2.png}
  \end{subfigure}
  \caption{\it Example of non-nested multigrid levels.}\label{fig:nonnested}
\end{figure}

The current implementation extends the existing global-coarsening infrastructure by
introducing, in addition to the (conformal) \texttt{MGTwoLevelTransfer},
 a new (non-conformal) two-level transfer operator:
\begin{c++}
MGTwoLevelTransferNonNested two_level_transfer(/*...*/);

two_level_transfer.reinit(dof_handler_fine, dof_handler_coarse,
                          mapping_fine, mapping_coarse,
                          constraints_fine, constraints_coarse)
\end{c++}
The resulting object can be passed to \texttt{MGTransferGlobalCoarsening} (and
\texttt{MGTransferBlockGlobal\allowbreak{}Coarsening}) in the usual way. The
interface of the new class is purposefully similar to
\texttt{MGTwoLevel\allowbreak{}Transfer},
with the difference that the former (i) also
takes \texttt{Mapping} instances, and (ii) is not limited to the case
that the coarser of the two \texttt{DoFHandler} instances can be generated by a coarsening step from the fine
mesh.

At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator provides a prolongation operator
that prolongates a vector from the coarse space to the fine space,
by a pointwise interpolation (injection) from the coarse mesh to the support points of the fine mesh. This is done
by looping over all coarse cells and interpolating to all (fine support) points that fall into the cell with
\texttt{FEPointEvaluation}. The local result is scattered into the fine vector, which is finalized by a
communication and a weighting step that is necessary as multiple elements could add to the same global entry.

The current implementation works for scalar, continuous (\texttt{FE\_Q})
and discontinuous (\texttt{FE\_DGQ}) elements on hypercube-shaped cells. We plan to support
vectorial and simplex elements and to add support
for projection, in addition to the pointwise interpolation outlined
above. Furthermore, we envision providing utility tools to generate
coarser
meshes, based on a fine mesh. Currently, the users themselves
have to provide such meshes,
e.g., by generating meshes with different cell sizes in external mesh-generation tools.

Note that the new class \texttt{MGTwoLevelTransferNonNested} is not limited to
multigrid but can be used to prolongate results and restrict residuals between any
two meshes. Indeed, the infrastructure has been successfully applied to perform
conservative interpolation between a fine (near) mesh and a coarse (far) mesh in the
context of aeroacoustic problems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New linear algebra features}\label{sec:lac}

There are a number of linear-algebra related new features in this release:
\begin{itemize}
  \item Both \texttt{SolverGMRES} and \texttt{SolverFGMRES} now support the classical
  Gram--Schmidt orthonormalization in addition to the modified Gram--Schmidt algorithm. This
  reduces the cost of vector operations in terms of
  communication latency and memory transfer significantly.
  \item Our Chebyshev preconditioner (\texttt{PreconditionChebyshev}) now also
  supports James Lottes’s novel fourth-kind Chebyshev
  polynomial~\cite{lottes2022optimal, phillips2022optimal}.
  \item Our relaxation preconditioner (\texttt{PreconditionRelaxation}) now also
  allows interleaving cell loops and vector updates related to relaxation. The
  relaxation iteration reads as
  \begin{align*}
  x^{(i+1)} \gets x^{(i)} + \omega P^{-1}(b-Ax^{(i)}).
  \end{align*}
  In the case that the preconditioner $P$ is a diagonal matrix, the zeroing of
  the destination vector $x_{i+1}$ can be performed during a
  \texttt{pre}-operation  of the application of $A$
  and the vector update $x^{(i+1)}_j \gets x^{(i)}_j + \omega P^{-1}_{j,j}(b_j-(Ax^{(i)})_j)$
  during a \texttt{post} operation, allowing for a reduction in the number of read and write
  accesses from 8 and 4, to 1 and 3, respectively. The existing \texttt{pre}/\texttt{post}
  support in our Chebyshev-preconditioner implementation (\texttt{PreconditionChebyshev})
  has been improved. In addition, both \texttt{PreconditionRelaxation} and
  \texttt{PreconditionChebyshev} support \texttt{pre}/\texttt{post} optimizations
  now not only for diagonal preconditioners but also for preconditioners that are
  built around cell loops and, as a consequence, support interleaving. Examples of
  such preconditioners are patch-based additive Schwarz preconditioners.
  \item In the context of additive Schwarz methods, the preconditioner application
  is defined as
  \begin{align*}
  v = P^{-1} u = \sum R_i^\top A_i^{-1} R_i u
  \end{align*}
  with $A_i = R_i A R_i^T$ being a block of the assembled system
  matrix $A$ restricted
  to an index set described by $R_i$. In the special case when $R_i$ denotes the unknowns
  of cells, the expression $R_i A R_i^T$ resembles the reverse of matrix
  assembly.
  During the restriction step,
  rows of the system matrix that are potentially owned by other
  processes are needed.  In \dealii, it is not possible to access remote entries
  of sparse matrices. Two new functions query this information. First,
  \texttt{restrict\_to\_serial\_sparse\_matrix()} creates, based
  on a given index set, a serial
  sparse matrix from a distributed matrix:
\begin{c++}
SparseMatrixTools::restrict_to_serial_sparse_matrix (
  sparse_matrix_in, sparsity_pattern, requested_index_set,
  system_matrix_out, sparsity_pattern_out)
\end{c++}
This function can be used, e.g., if the granularity of the additive Schwarz preconditioner
is a complete subdomain, potentially, with a fixed overlap.

In contrast, \texttt{restrict\_to\_full\_matrices()} performs the restriction
for arbitrary number of patches/blocks:

\begin{c++}
SparseMatrixTools::restrict_to_full_matrices (
  sparse_matrix_in, sparsity_pattern, indices_of_blocks, blocks)
\end{c++}

The data is stored in dense matrices, since the typical granularity is a (rather small) cell-centric or
vertex-star patch.

>>>>

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Marc Fehling,
    Johannes Heinz,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.6, 2024},
}

\title{The \dealii Library, Version 9.6}

@@

\author[4]{Marc~Fehling}
 \affil[4]{Charles University, Czech Republic}

\author[5]{Timo~Heister}
 \affil[5]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister@clemson.edu}}}

\author[6]{Luca~Heltai}
\affil[6]{University of Pisa, Italy}

 \author[7]{Martin~Kronbichler}
 \affil[7]{Ruhr University Bochum, Germany}

\author[8]{Matthias~Maier}
\affil[8]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[9]{Peter Munch}
 \affil[9]{Uppsala University, Sweden.
  {\texttt{peter.munch@it.uu.se}}}

@@

\author[10]{David Wells}
\affil[10]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.6.
\end{abstract}

@@

\dealii version 9.6.0 was released X X, 2024.
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii software library version 9.6. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item 
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item changes of interface of AffineConstraints
  \item clean up: orientation
\end{itemize}
%
The changelog lists more than X other features and bugfixes.

@@

\begin{itemize}
\item Hermite element
\item simplices: cubic simplex elements, improved refinement of tetrahedron cells
\item Tpetra advances
\item Nedelec for meshes with hanging nodes
\end{itemize}

@@

\begin{itemize}
\item VectorizedArray for ARM Neon
\end{itemize}

@@

\begin{itemize}
\item non-nested multigrid: new features and performance improvement
\item FERemoteEvaluation
\item FEColungValues
\end{itemize}
